\ttitle

### 2013-06-04 Chapters/Chapter2.tex
Rough finish of chapter 2

%----------------------------------------------------------------------------------------
\section{Outline}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction)
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

@@

*** Remove notion of GLR ***

The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
The former originates from control methods for detection from bench marks.
The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
These two methods is discussed and analyzed in section \ref{cusum}.

The GLR method, as with others, relies on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the PDF, instead of estimating each PDF.
This approach is discussed and analyzed in section \ref{density-ratio}.

The density-estimation methods, as with the GLR, rely on the log-likelihood ratio between PDFs.
The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a SVM (Support Vector Machine).
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using kernel trick.
A change in the PDF is represented by a change in the radius of this sphere.
Section \ref{svm} discusses the SVM-method.

@@

>>>>

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction)
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

@@

The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection} takes this approach.
It originates from control methods for detection from bench marks.
This method and some derivatiges are discussed and analyzed in section \ref{cusum}.

Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the PDF, instead of estimating each PDF.
This approach is discussed and analyzed in section \ref{density-ratio}.

The density-estimation methods rely on the log-likelihood ratio between PDFs.
The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a SVM (Support Vector Machine).
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using kernel trick.
A change in the PDF is represented by a change in the radius of this sphere.
Section \ref{svm} discusses the SVM-method.

@@

*** Maybe add dimensionality reduction, but for now leave out ***

### 2013-06-04 Chapters/Chapter2/cusum.tex
Rough finish of chapter 2

*** read book Basseville \cite{basseville1993detection} ***

Papers:
\begin{itemize}
  \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
  Implemented this algorithm.
  \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
  \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
\end{itemize}

\subsection{CUSUM}
Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
The CUSUM (cumulative sum) method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
Primary for detection of mean shift.
The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}.
Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}

An other used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the Moving Sum (MOSUM) of squares by \cite{hsu2007mosum}.

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of CUSUM to detect changes in variance.
Using an iterative method algorithm --- Iterative Cumulative Sum of Squares (ICSS), they are able to find multiple change points in a reflective manner.
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum is squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
When there is a sudden change, the value will increase and exceed some predefined boundary with high probability.
The behavior of $D_k$ is related a Brownian bridge.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

>>>>

% *** read book Basseville \cite{basseville1993detection} ***

% Papers:
% \begin{itemize}
%   \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
%   Implemented this algorithm.
%   \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
%   \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
% \end{itemize}

Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
The CUSUM (cumulative sum) method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
Primary for detection of mean shift.
The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}.
Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}

An often used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
It is an non-Bayesian method and thus makes no assumptions (or: prior belief distributions) for the change points.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the Moving Sum (MOSUM) of squares by \cite{hsu2007mosum}.

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of CUSUM \cite{brown1975techniques} to detect changes in variance.
Using the Iterative Cumulative Sum of Squares (ICSS) algorithm they are able to find multiple change points in a reflective manner.
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum is squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
When there is a sudden change, the value will increase and exceed some predefined boundary with high probability.
The behavior of $D_k$ is related a Brownian bridge.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

### 2013-06-04 Chapters/Chapter2/density_ratio.tex
Rough finish of chapter 2

Papers:
\begin{itemize}
  \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
  \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
  \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
  \item Direct importance estimation for covariate shift adaptation, 2008, 84 refs \cite{sugiyama2008direct}
\end{itemize}

@@

Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
A limitation of these methods is that they rely on pre-specified parametric models.
Non-parametric models for density estimation have been proposed, but it is said to be a hard problem.
A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
One of the recent methods to achieve this is the Kullback-Leibler Importance Estimation Procedure (KLIEP) by Sugiyama \etal \cite{sugiyama2008direct}.

The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the KLIEP algorithm.
The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
An advantage over other non-parametric approaches, such as sequential one-class support vector machines is that the model has an natural cross-validation procedure.
This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.

>>>>

% Papers:
% \begin{itemize}
%   \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
%   \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
%   \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
%   \item Direct importance estimation for covariate shift adaptation, 2008, 84 refs \cite{sugiyama2008direct}
% \end{itemize}

@@

Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
A limitation of these methods is that they rely on pre-specified parametric models.
Non-parametric models for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
One of the recent methods to achieve this is the Kullback-Leibler Importance Estimation Procedure (KLIEP) by Sugiyama \etal \cite{sugiyama2008direct}.

The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the KLIEP algorithm.
The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
An advantage over other non-parametric approaches, such as sequential one-class support vector machines, is that the model has an natural cross-validation procedure.
This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.

### 2013-06-04 Chapters/Chapter2/svm.tex
Rough finish of chapter 2

Introduced by Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
This process is referred to as the kernel trick.

The method of Camci \cite{camci2010change} uses one-class support vector machines to segment time series data.
For the current interval a hyper-sphere is constructed, using the kernel trick, which circumscribes all the data.
To cope with possible errors or outliers a soft-margin is applied.
The data can thus be represented using a center $c_1$ and radius $r_1$ of the hyper-sphere.
New data points are consecutively mapped from the input space to the feature space.
The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.

\subsection{One-class Support Vector Machine}
*** This section is too detailed for the Literature Review, but keep it for now ***

*** Replace with / add information from summary in \ref{svm-explained}. ***

The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
\begin{equation}
  \mathrm{Min}\ r^2
\end{equation}
\begin{equation}
  \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
\end{equation}

@@

*** Add new function and constraints? ***

@@

*** First tell more about (underlying) probability density functions, to relate to other methods ***

In case the second window of data points does not belong to the same class, i.e. the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
The amount of difference can be expressed by a dissimilarity measure over the representations.
When the dissimilarity between the two windows exceeds some predefined threshold $th$, there exists a change point between the windows.

This process can be visualized as done in *** insert figure of four circles ***. The second window, $h_2$ can be constructed from the first by e.g. a shift of one data point. *** explain data point positions by circle ***.

Note that a difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.

>>>>

Introduced by Cortes and Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
This process is referred to as the kernel trick.

The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
To cope with possible errors or outliers a soft-margin is applied \cite{cortes1995support}.
The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
New data points are consecutively mapped from the input space to the feature space.
The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
\begin{equation}
  \mathrm{Min}\ r^2
\end{equation}
\begin{equation}
  \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
\end{equation}

@@

@@

In case the second window of data points does not belong to the same class, i.e. the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
A difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.
The amount of difference can be expressed by a dissimilarity measure over the representations.
When the dissimilarity between the two windows exceeds some predefined threshold $th$, there exists a change point between the windows.

*** Give visual explanation with circles and in- and out-side new data points ***
